### 2022-03-02 paper3.tex
Updates

TODO
Mirbabaie et al. (2020)
@ TERNULLO

@@

After spreading around the world in a matter of months, the coronavirus
(COVID-19) became a leading cause of death in the United States. Although the
Centers for Disease Control and Prevention(CDC)
(2020) proposed several potential
mitigation strategies, the method of mitigation that received the most national
attention is the call to stay-at-home and social distance for non-essential
workers. CDC officials and frontline health care professionals advise that the
best way to prevent exposure to the virus is to stay at home and avoid close
contact with people.

@@

Although these public health interventions were pushed by policy makers
throughout 2020 and 2021, policy makers often struggle to promote public
adoption of policies that rely on the publics' willingness to adjust their
current `risky' behaviors to less-risk behaviors based on scientific facts. The
lack of adaption to newly recommended behaviors is partially due to the
variation in how individuals came to understand the threats of the disease
(Akpan et al. 2021; Bailey et al. 2020), but also due to habitus: the
ingrained habits, skills and dispositions of individuals and groups
(Bourdieu 1977). Habitus shapes how individuals perceive different social
interactions and norms and shape how they adopt different policies, leading to
different social outcomes and a divergence of attitudes (Madeira et al. 2018; Scott-Arthur, Brown, and Saukko 2021; Williams 1995).

There is little research on stay-at-home rates, especially in cases of pandemics
or disasters. Though research on the causes of population mobility did expand
after the onset of the Covid-19 pandemic (Bargain and Aminjonov 2020; Bourassa 2020; Bourassa et al. 2020; Grossman et al. 2020; Hagger et al. 2020; Hill, Gonzalez, and Burdette 2020; Hill, Gonzalez, and Davis 2021; Hill, Gonzalez, and Upenieks 2021; Huynh 2020), most research focused on unchanging cultural
determinants of the reduction in stay-at-home rates. For instance,
Gibbons, Yang, and Oren (2021) finds that social capital had spatially inconsistent effects on
stay-at-home rates. Instead of focusing on cultural relationships with
stay-at-home rates, this paper will instead focus on the dynamic nature of
population mobility across population networks.

There is more research on vaccine uptake
(Schmid et al. 2017) because of increased hesitancy and
anti-vaccination movements over previous decades
(Baumgaertner, Carlisle, and Justwan 2018; Hornsey et al. 2020; Johnson et al. 2020; Whitehead and Perry 2020). While some studies use social network analysis to
study the discussions for an against vaccinations online
(Milani, Weitkamp, and Webb 2020), vaccination rates are a great way to
investigate the establishment of new social norms. First, vaccination uptake is
measurable with stable information tracking infrastructure and vaccination
rates are varied across locations and time. Secondly, because vaccines have
become politicized (Motta 2021; Motta et al. 2021), the social contagion effects can be
studied.

@@

Individuals engage with each other and their distributive ties to create
community contexts where norms, beliefs, and values circulate. These clusters of
interaction are called social networks, and if ``each person continues to
interact primarily with others nearby in space, the forces of conformity will be
strongest locally, leading to the emergence of clusters of people sharing
similar behavior'' (Kitts and Shi 2018). This community
interaction ultimately leads to converged communities of belief structures with
variations in how much they diverge from the norm (Cullum and Harton 2007; Latané and Bourgeois 1996; Okada 2017).

``Culture'' diffuses through communities and social networks. Information and
opinions spread (Robert M. Bond et al. 2012b; Fowler and Christakis 2010; Klar and Shmargad 2017), behaviors are adopted
(Aral and Nicolaides 2017; Centola 2010, 2011; Christakis and Fowler 2008; Rosenquist et al. 2010) , and there are patterns of health
contagion (Cacioppo, Fowler, and Christakis 2009; Christakis and Fowler 2007). However,
``different things spread in different ways and to different extents''
(Christakis and Fowler 2013, p 563) and when modeling diffusion and
contagion, we must be very specific about our scope conditions as they are
relevant to our theory and not to cross theories to infer connections where they
may not exist (Kitts and Quintane 2020).

Most of the diffusion literature does not focus on establishment of new norms
but the adoption of culture and specific deviant behaviors (see
Centola and Baronchelli (2015) for an exception). DellaPosta and
colleagues (2015) outline how the spread of culture
and behavior is tied to network autocorrelation, or ``the tendency for people to
resemble their network neighbors.'' They show that the distance between to agents
in sociocultural space can determine the likelihood of the adoption of a new
behavior. Like Axelrod (1997), this outlines
how the local convergence of close network actors becomes amplified and can
lead to global polarization between groups.

Outdated models found in early public health research claim that information
about the risks of behaviors will lead to changes in behaviors to mitigate those
said risks (e.g. Flay, DiTecco, and Schlegel 1980). However, while this model can be valid in specific
cases, more research has shown how the risks themselves do little to motivate
behavioral change (Witte and Allen 2000; Wolburg 2006). Often, an appeal to fear is
found to be a major driver of the adoption of information campaigns and some
research has shown that ``social network exposure to COVID-19 cases shapes
individuals'' beliefs and behaviors concerning the coronavirus''
(Bailey et al. 2020). Because of this, it is logical that higher local incidences of
infection would inspire fear and increase adherence to public health measures
according to this older model.

@@

Centola and Macy (2007) theorize that simple contagions are not
enough to spread behavioral change. Simple contagions are those in which only
one point of contact is needed to receive contagion, like with infectious
disease or to learn simple bits of information.
Centola and Macy (2007)'s large contribution was the theorizing
of complex contagions, those that require ``independent affirmation or
reinforcement from multiple sources'' (p 703) and is not based on the number of
exposures but the number of sources of exposure. In the case of behavioral
contagion, this means that the behavior must be reinforced through witnessing
multiple alters perform this behavior before contagion can take effect.

@@

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{G:/My Drive/R/My Projects/dissertation/figure/paper3/dag}

}

\caption{Elaboratory Theoretical Model of Health Behavior Norms}\label{fig:dag}
\end{figure}

@@

While much of the social contagion literature, like the theories above, focuses
on structural boundaries and homophily as causes of how diffusion occurs,
Goldberg and Stein (2018) propose a
disrupting alternative mechanism. They argue that what actually diffuses during
social contagion are the perceptions about which beliefs or behaviors are
compatible with one another, what they call ``associative diffusion.'' This
argument that culture does not spread like a virus but instead is dependent on
how belief structures are connected to each other is important to test because
norms around health behaviors became politicized issues during the Covid-19
pandemic. This means that the stay-at-home or vaccination behavior themselves
were not contagious, but the cognitive association of precisely what social
distancing or vaccination uptake \emph{meant} were spreading between populations of
individuals. Moreover, as Houghton (2021) outlines, diffusants - the things that are
diffusing through a population, are not independent of each other
(Mason, Conrey, and Smith 2007). When we take into account the interdependence between different
beliefs that are diffusing, such as Covid-19 is a hoax, social distancing saves
lives, the Covid-19 vaccine includes a microchip to control your thoughts, and lockdown
is against the rights guaranteed in the US constitution, we can imagine how
these various diffusants become associated together into structures of belief
through schemas of perception (Houghton 2021). While this cognitive theory of
cultural variation is difficult to test, the theory it supplies provides a solid
framework for how behavioral norms formed during the pandemic.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{G:/My Drive/R/My Projects/dissertation/figure/paper3/model}

}

\caption{Preliminary model of norm formation}\label{fig:norm-framework}
\end{figure}

The figure above illustrates how I interpret associative diffusion will impact
stay-at-home and vaccination rates in times of unsettled norms. An
individual enters this framework when they realize they don't have a normative
set of responses in their cultural toolkit to respond to an unfamiliar situation
they are presented with. Individuals look to norms to regulate behavior, avoid
deviance, and to maintain order (Horne and Mollborn 2020; Shepherd 2017); when they don't have a normative behavior
to follow (or rebel against), individuals look to the other sources in their
``community'' to mimic behavior, such as high-status individuals, institutions,
and members of their social network. I theorize that individuals look towards
their physical community, to their social network, popular media (which may
include government and science recommendations), established norms that they may
find online through search, and to the real threat of infection (what would
happen if I do nothing about this norm). Following both the Integrated
Theoretical Framework of Norms (Horne and Mollborn 2020) and
associative diffusion (DellaPosta et al. 2015; Goldberg and Stein 2018), I theorize that individuals interpret
signals from their ``community'' through their cognitive biases and behavioral
predispositions to determine their formation of new behavioral norms.

There is some evidence I can draw upon to support the associative diffusion
fraemwork. For instance, there is a central conflict between religious and
scientific ideologies which I theorize leads more religious counties to reject
the stay-at-home order and vaccines (Evans and Evans 2008)
because the associative diffusion of these behaviors pits public health
recommendations against religious ideologies.
Furthermore, as COVID-19 has become a politically polarizing issue, conservative
ideology and a general mistrust of ``big government'' (Frank 2007; Gauchat 2012) are likely to lead to resistance to the
government and scientific guidance.

@@

\begin{table}[!h]

\caption{\label{tab:google-desc-table}Descriptive Statistics for Stay-at-Home Models (county-level)}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
  & min & max & mean & sd\\
\midrule
Percent White & \num{0.14} & \num{0.98} & \num{0.81} & \num{0.15}\\
Percent College Graduates & \num{0.08} & \num{0.75} & \num{0.26} & \num{0.10}\\
Percent over 65 & \num{6.60} & \num{56.70} & \num{16.90} & \num{4.05}\\
Median Income & \num{28951.00} & \num{142299.00} & \num{59773.61} & \num{15264.50}\\
Monthly Unemployment Rate & \num{1.40} & \num{34.60} & \num{8.12} & \num{4.00}\\
Percent of GOP votes, 2016 & \num{0.08} & \num{0.90} & \num{0.57} & \num{0.15}\\
Percent Evangelical Christian & \num{0.00} & \num{0.67} & \num{0.20} & \num{0.14}\\
'Fox News' Trend & \num{0.00} & \num{204.08} & \num{29.59} & \num{16.94}\\
'Social Distancing' Trend & \num{0.00} & \num{500.00} & \num{20.64} & \num{25.36}\\
'Covid Conspiracy' Trend & \num{0.00} & \num{825.00} & \num{42.51} & \num{80.44}\\
Covid Case Rate & \num{0.00} & \num{343.71} & \num{21.17} & \num{27.44}\\
Week Number & \num{0.00} & \num{43.00} & \num{23.14} & \num{11.77}\\
Movement Signal & \num{-2.06} & \num{26.44} & \num{8.31} & \num{4.22}\\
Movement Discordance & \num{0.15} & \num{11.05} & \num{2.22} & \num{0.80}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}Raw values presented in table. Values in models are normalized.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}Notes: 1,375 counties, March 02 through December 28, 2020.}\\
\end{tabular}
\end{table}
\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{dissertation_files/figure-latex/plot-google-1}

}

\caption{Stay at Home Rates over Time}\label{fig:plot-google}
\end{figure}

My first dependent variable aims to operationalize behavioral norms as
stay-at-home rates with data from Google (Google 2020). While the Google
COVID-19 Community Mobility Reports include multiple measures of mobility based
on location and activity information, the change in time spent in residence
most closely aligns to an operationalization of an obedience to CDC and governor
orders, a new and emerging norm in 2020. The change in time spent in residence
variable represents percent change of time spent from Google Location History
within geographic areas that Google has designated as a residential area. These
data are aggregated to the county-level based on anonymized sets of data from
users who have turned on the Location History setting, which is off by default.
This means our sample is possibly skewed to people in the United States a) with
a mobile phone, b) with a Google account, and c) with knowledge of how to
synchronize their location history. Google has not made it clear whether there
are any weights in place to correct for these potential biases. Google
calculates the relative change in mobility in comparison to the median value of
movement in the area for the same the corresponding day of the week, during the
5-week period Jan 3--Feb 6, 2020. As Google did not provide data on certain US
counties if they did not have sufficient statistically significant levels of
data, my final county sample is \emph{n} = 1375
(compared to the total population of 3,107 US counties). Particular areas that
are excluded in this analysis include all counties in Alaska and DC, and over
half of the counties in North Dakota, South Dakota. A full list of excluded
counties available upon request. These data are longitudinal measures calculated
by creating a moving average of stay-at-home rates for each county (rolling
window width = 7 days) and then sampling every
Monday in the sample for 44 measures from March 02, 2020 through December 28,
2020.

@@

\begin{table}[!h]

\caption{\label{tab:vacc-desc-table}Descriptive Statistics for Vaccination Models (county-level)}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
  & min & max & mean & sd\\
\midrule
Percent White & \num{0.08} & \num{1.00} & \num{0.83} & \num{0.17}\\
Percent College Graduates & \num{0.03} & \num{0.78} & \num{0.22} & \num{0.10}\\
Percent over 65 & \num{3.20} & \num{56.70} & \num{18.86} & \num{4.50}\\
Median Income & \num{21504.00} & \num{142299.00} & \num{53522.45} & \num{14312.25}\\
Monthly Unemployment Rate & \num{0.90} & \num{22.00} & \num{4.96} & \num{1.95}\\
Percent of GOP votes, 2020 & \num{0.09} & \num{0.94} & \num{0.64} & \num{0.16}\\
Percent Evangelical Christian & \num{0.00} & \num{1.31} & \num{0.22} & \num{0.16}\\
'Fox News' Trend & \num{5.00} & \num{405.88} & \num{72.12} & \num{27.38}\\
'Covid-19 vaccine' Trend & \num{0.00} & \num{755.56} & \num{74.67} & \num{70.33}\\
'Covid Conspiracy' Trend & \num{0.00} & \num{2000.00} & \num{56.05} & \num{189.61}\\
Covid Case Rate & \num{0.00} & \num{998.58} & \num{22.79} & \num{27.23}\\
Week Number & \num{1.00} & \num{35.00} & \num{18.00} & \num{10.10}\\
Vaccination Signal & \num{0.00} & \num{62.84} & \num{19.62} & \num{14.84}\\
Vaccination Discordance & \num{0.00} & \num{22.40} & \num{5.26} & \num{3.62}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}Raw values presented in table. Values in Models are normalized.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}Notes: 2,819 counties, January 04 through August 30, 2021.}\\
\end{tabular}
\end{table}
\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{dissertation_files/figure-latex/plot-vacc-1}

}

\caption{Vaccination Rates over Time}\label{fig:plot-vacc}
\end{figure}

The second dependent variable aims to operationalize vaccination uptake through
vaccination rate information in the United States from January 2021 through
August 30 2021 (CDC 2021). Data represents all vaccine partners including
jurisdictional partner clinics, retail pharmacies, long-term care facilities,
dialysis centers, Federal Emergency Management Agency and Health Resources and
Services Administration partner sites, and federal entity facilities.
Vaccination data is available for all US counties with the exception of parts of
California and Massachusetts, Hawaii, and Texas. In Texas and Hawaii, no county
level information is available, and California does not report the county of
residence for vaccinations when the county of residence has a population less
than 20,000 people. Finally, Massachusetts does not provide vaccination data for
Barnstable, Dukes, and Nantucket counties because of their small populations.
Therefore, my final county sample for vaccination rates is \emph{n} = 2819
(compared to the total population of 3,107 US counties). A list of included FIPS
are available upon request. Vaccination rates are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. I
sample every Monday between January 2021 through August 30, 2021 for 35 total
observations for each county.

@@

I utilize the likelihood of a friendship connection between counties
to create a county-level social network weighted by the probability of a tie.
Using this network and the two independent variables above, I examine how the
vaccination and stay-at-home rates of peer counties is contagious to the
ego-county. To do this, I first take the weighted average of each ego's network
signals using equation \eqref{eq:networksignal} where \texttt{x} denotes the
vaccination and stay-at-home rates of each alter county and \texttt{w} represents the
likelihood of a friendship connection between the ego county and each of their
alters, lagged by one week. This `signal' of norms gives us insight into the
coevolving contagion patterns of the new social norm being established. The
likelihood of friendship connection comes from the ``Social Connectedness Index''
(Bailey, Rachel Cao, et al. 2018; Facebook 2020) which indexes the social links between geographies by
the likelihood of Facebook friends. It is an aggregated measure of Facebook
friendship connections between counties. It corresponds to ``the (relative)
probability that two arbitrary Facebook users across two geographies are friends
with each other.'' The data is available at various geographic aggregation
levels, such as U.S. zip codes or entire countries. However, data is only
available at one time point because Facebook has found that although there are
individual changes in friendship connections over time, the aggregate
statistical probabilities of friendship remain stable. This measure is based on
user-provided information and real-time location-service data gathered by
Facebook. Facebook data has been shown to be highly representative of the U.S.
population and Facebook friendship links largely represent real-world
friendships (Bailey, Ruiqing Cao, et al. 2018; Jones et al. 2013). The data has been tested to show
how initial COVID-19 hotspots are related to subsequent virus spread to
non-hotspots, even after controlling for population density and geographic
distance (Kuchler, Russel, and Stroebel 2020).

@@

Figures \ref{fig:assortativity-vacc} and \ref{fig:discordancenetwork} provide
a visualization of both network signal and signal discordance for April 26, 2021
for 7 selected counties.\footnote{Figures in this paper were all created using ggplot2
  (Wickham 2011; Wickham et al. n.d.), patchwork (Pedersen n.d.), and ggridges (Wilke n.d.)}
While a county like Lake County, Ohio has low discordance meaning the signal of
vaccination rate is reinforced, Navajo County, Arizona receives very diverse
signals from their county-alters, negating any contagion effects. Figure
\ref{fig:discordancenetwork} explicitly outline how the weights and signals are
calculated for an ego with only 5 alters.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{dissertation_files/figure-latex/assortativity-vacc-1}

}

\caption{How Signal Mean and Signal Assortativity are Measured}\label{fig:assortativity-vacc}
\end{figure}

\begin{figure}

\begin{center}\includegraphics[width=0.5\linewidth,]{dissertation_files/figure-latex/discordancenetwork-1} \end{center}
  \begin{equation}
    w = \begin{bmatrix}0.0000015\\0.0000023\\0.0000527\\0.0000966\\0.0001119\end{bmatrix},  
    x = \begin{bmatrix}29.9\\20.4\\27.7\\24.6\\24.3\end{bmatrix} \longrightarrow
    \begin{matrix} \bar{x}, signal = 25.083\\ s, discordance = 1.41 \end{matrix} 
  \end{equation}
  \caption{Visualization of how Network Signal and Discordance are Calculated}
  \label{fig:discordancenetwork}
\end{figure}

@@

To estimate the concept of `real threat of infection' in figure
\ref{fig:norm-framework}, the models use county-level Covid-19 case rates with
data compiled by The New York Times (2022). It is widely acknowledged that there are biases in
this data due to inconsistencies and availability in testing as well as
different community propensity to test (CDC 2020b; Gu n.d.) However, it is the best
measure we have of actual case rates. County data are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. Case
Rate is measured as number of cases per 100,000 population. Observations vary
from a minimum of 0 to a maximum of 1,565 across the two datasets.

@@

\begin{table}[!h]

>>>>

% TODO
\citep{mirbabaie_etal20, ternullo22}

@@

After spreading around the world in a matter of months, the coronavirus
(COVID-19) became a leading cause of death in the United States. Although the
Centers for Disease Control and Prevention (CDC) \citeyearpar{centersfordiseasecontrolandpreventionHowProtectYourself2020} proposed several potential mitigation strategies, the method of mitigation that received the most national
attention is the call to stay-at-home and social distance for non-essential
workers. CDC officials and frontline health care professionals advise that the
best way to prevent exposure to the virus is to stay at home and avoid close
contact with people.

@@

Although these public health interventions were pushed by policy makers
throughout 2020 and 2021, policy makers often struggle to promote public
adoption of policies that rely on the publics' willingness to adjust their
current `risky' behaviors to less-risk behaviors based on scientific facts. The
lack of adaption to newly recommended behaviors is partially due to the
variation in how individuals came to understand the threats of the disease
\citep{akpanAssociationWhatPeople2021, bailey_etal20}, but also due to habitus: the
ingrained habits, skills and dispositions of individuals and groups
\citep{bourdieu77}. Habitus shapes how individuals perceive different social
interactions and norms and shape how they adopt different policies, leading to
different social outcomes and a divergence of attitudes \citep{scottarthur_etal21, williams95, madeira_etal18}.

There is little research on stay-at-home rates, especially in cases of pandemics
or disasters. Though research on the causes of population mobility did expand
after the onset of the Covid-19 pandemic 
\citep{bargainTrustCompliancePublic2020, bourassaStatelevelStayathomeOrders2020, bourassaSocialDistancingHealth2020, grossmanPoliticalPartisanshipInfluences2020, haggerPredictingSocialDistancing2020, hillBloodChristCompels2020, hillNastiestQuestion, hillLoveThyAged2021,huynhDoesCultureMatter2020}, most research focused on unchanging cultural
determinants of the reduction in stay-at-home rates. For instance, \citet{gibbons_etal21}
 finds that social capital had spatially inconsistent effects on
stay-at-home rates. Instead of focusing on cultural relationships with
stay-at-home rates, this paper will instead focus on the dynamic nature of
population mobility across population networks.

There is more research on vaccine uptake \citep{schmidBarriersInfluenzaVaccination2017}
 because of increased hesitancy and
anti-vaccination movements over previous decades \citep{baumgaertnerInfluencePoliticalIdeology2018, hornseyDonaldTrumpVaccination2020, johnsonOnlineCompetitionPro2020, whiteheadHowCultureWars2020}. While some studies use social network analysis to
study the discussions for an against vaccinations online \citep{milaniVisualVaccineDebate2020}, vaccination rates are a great way to
investigate the establishment of new social norms. First, vaccination uptake is
measurable with stable information tracking infrastructure and vaccination
rates are varied across locations and time. Secondly, because vaccines have 
become politicized \citep{mottaRepublicansNotDemocrats2021, mottaIdentifyingPrevalenceCorrelates2021}, the social contagion effects can be
studied.

@@

Individuals engage with each other and their distributive ties to create
community contexts where norms, beliefs, and values circulate. These clusters of
interaction are called social networks, and if ``each person continues to
interact primarily with others nearby in space, the forces of conformity will be
strongest locally, leading to the emergence of clusters of people sharing
similar behavior'' \citep{kitts_shi18}. This community
interaction ultimately leads to converged communities of belief structures with
variations in how much they diverge from the norm \citep{cullumCulturalEvolutionInterpersonal2007, lataneExperimentalEvidenceDynamic1996, okadaStructureCulturalRejection2017}.

``Culture'' diffuses through communities and social networks. Information and
opinions spread \citep{bond_etal12, fowler2010cooperative,klarEffectNetworkStructure2017}, behaviors are adopted \citep{aralExerciseContagionGlobal2017, centolaSpreadBehaviorOnline2010, centolaExperimentalStudyHomophily2011,christakis2008collective,rosenquist2010spread}, and there are patterns of health contagion \citep{cacioppo2009alone,christakisSpreadObesityLarge2007}. However,
``different things spread in different ways and to different extents''
\citep[p. 563]{christakisSocialContagionTheory2013} and when modeling diffusion and
contagion, we must be very specific about our scope conditions as they are
relevant to our theory and not to cross theories to infer connections where they
may not exist \citep{kitts_quintane20}.

Most of the diffusion literature does not focus on establishment of new norms
but the adoption of culture and specific deviant behaviors 
\citep[see][for an exception]{centolaSpontaneousEmergenceConventions2015}. 
\citet{dellapostaWhyLiberalsDrink2015} outline how the spread of culture
and behavior is tied to network autocorrelation, or ``the tendency for people to
resemble their network neighbors.'' They show that the distance between to agents
in sociocultural space can determine the likelihood of the adoption of a new
behavior. Like \citet{axelrodDisseminationCultureModel1997}, this outlines
how the local convergence of close network actors becomes amplified and can
lead to global polarization between groups.

Outdated models found in early public health research claim that information
about the risks of behaviors will lead to changes in behaviors to mitigate those
said risks \citep[e.g.][]{flay_etal80}. However, while this model can be valid in specific
cases, more research has shown how the risks themselves do little to motivate
behavioral change \citep{witte_allen00,wolburg06}. Often, an appeal to fear is
found to be a major driver of the adoption of information campaigns and some
research has shown that ``social network exposure to COVID-19 cases shapes
individuals'' beliefs and behaviors concerning the coronavirus'' \citep{bailey_etal20}. Because of this, it is logical that higher local incidences of
infection would inspire fear and increase adherence to public health measures
according to this older model.

@@

\citep{centolaComplexContagionsWeakness2007} theorize that simple contagions are not
enough to spread behavioral change. Simple contagions are those in which only
one point of contact is needed to receive contagion, like with infectious
disease or to learn simple bits of information.
Centola and Macy \citeyearpar{centolaComplexContagionsWeakness2007}'s large contribution was the theorizing
of complex contagions, those that require ``independent affirmation or
reinforcement from multiple sources'' (p. 703) and is not based on the number of
exposures but the number of sources of exposure. In the case of behavioral
contagion, this means that the behavior must be reinforced through witnessing
multiple alters perform this behavior before contagion can take effect.

@@

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper3/dag}}
\caption{Elaboratory Theoretical Model of Health Behavior Norms}\label{fig:dag}
\end{figure}

@@

While much of the social contagion literature, like the theories above, focuses
on structural boundaries and homophily as causes of how diffusion occurs,
\citet{goldbergSocialContagionAssociative2018} propose a
disrupting alternative mechanism. They argue that what actually diffuses during
social contagion are the perceptions about which beliefs or behaviors are
compatible with one another, what they call ``associative diffusion.'' This
argument that culture does not spread like a virus but instead is dependent on
how belief structures are connected to each other is important to test because
norms around health behaviors became politicized issues during the Covid-19
pandemic. This means that the stay-at-home or vaccination behavior themselves
were not contagious, but the cognitive association of precisely what social
distancing or vaccination uptake \emph{meant} were spreading between populations of
individuals. Moreover, as \citet{houghtonSocialContagionInteracting2021} outlines, diffusants - the things that are
diffusing through a population, are not independent of each other \citep{mason_etal07}. When we take into account the interdependence between different beliefs that are diffusing, such as Covid-19 is a hoax, social distancing saves
lives, the Covid-19 vaccine includes a microchip to control your thoughts, and lockdown
is against the rights guaranteed in the US constitution, we can imagine how
these various diffusants become associated together into structures of belief
through schemas of perception \citep{houghtonSocialContagionInteracting2021}. While this cognitive theory of cultural variation is difficult to test, the theory it supplies provides a solid framework for how behavioral norms formed during the pandemic.

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper3/model}}
\caption{Preliminary model of norm formation}\label{fig:norm-framework}
\end{figure}

The figure above illustrates how I interpret associative diffusion will impact
stay-at-home and vaccination rates in times of unsettled norms. An
individual enters this framework when they realize they don't have a normative
set of responses in their cultural toolkit to respond to an unfamiliar situation
they are presented with. Individuals look to norms to regulate behavior, avoid
deviance, and to maintain order \citep{horneNormsIntegratedFramework2020, shepherdStructurePerceptionHow2017}; when they don't have a normative behavior
to follow (or rebel against), individuals look to the other sources in their
``community'' to mimic behavior, such as high-status individuals, institutions,
and members of their social network. I theorize that individuals look towards
their physical community, to their social network, popular media (which may
include government and science recommendations), established norms that they may
find online through search, and to the real threat of infection (what would
happen if I do nothing about this norm). Following both the Integrated
Theoretical Framework of Norms \citep{horneNormsIntegratedFramework2020} and
associative diffusion \citep{dellapostaWhyLiberalsDrink2015, goldbergSocialContagionAssociative2018}, I theorize that individuals interpret
signals from their ``community'' through their cognitive biases and behavioral
predispositions to determine their formation of new behavioral norms.

There is some evidence I can draw upon to support the associative diffusion
fraemwork. For instance, there is a central conflict between religious and
scientific ideologies which I theorize leads more religious counties to reject
the stay-at-home order and vaccines \citep{evansReligionScienceEpistemological2008}
because the associative diffusion of these behaviors pits public health
recommendations against religious ideologies.
Furthermore, as COVID-19 has become a politically polarizing issue, conservative
ideology and a general mistrust of ``big government'' \citep{frank2007, gauchat2008} are likely to lead to resistance to the
government and scientific guidance.

@@

\input{tables/paper3/google-desc-table}

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper3/plot-google-1}}
\caption{Stay at Home Rates over Time}\label{fig:plot-google}
\end{figure}

My first dependent variable aims to operationalize behavioral norms as
stay-at-home rates with data from Google \citep{google2020}. While the Google
COVID-19 Community Mobility Reports include multiple measures of mobility based
on location and activity information, the change in time spent in residence
most closely aligns to an operationalization of an obedience to CDC and governor
orders, a new and emerging norm in 2020. The change in time spent in residence
variable represents percent change of time spent from Google Location History
within geographic areas that Google has designated as a residential area. These
data are aggregated to the county-level based on anonymized sets of data from
users who have turned on the Location History setting, which is off by default.
This means our sample is possibly skewed to people in the United States a) with
a mobile phone, b) with a Google account, and c) with knowledge of how to
synchronize their location history. Google has not made it clear whether there
are any weights in place to correct for these potential biases. Google
calculates the relative change in mobility in comparison to the median value of
movement in the area for the same the corresponding day of the week, during the
5-week period Jan 3--Feb 6, 2020. As Google did not provide data on certain US
counties if they did not have sufficient statistically significant levels of
data, my final county sample is \emph{n} = 1375
(compared to the total population of 3,107 US counties). Particular areas that
are excluded in this analysis include all counties in Alaska and DC, and over
half of the counties in North Dakota, South Dakota. A full list of excluded
counties available upon request. These data are longitudinal measures calculated
by creating a moving average of stay-at-home rates for each county (rolling
window width = 7 days) and then sampling every
Monday in the sample for 44 measures from March 02, 2020 through December 28,
2020.

@@

\input{tables/paper3/vacc-desc-table}

\begin{figure}
{\centering \includegraphics[width=0.8\linewidth]{figs/paper3/plot-vacc-1}}
\caption{Vaccination Rates over Time}\label{fig:plot-vacc}
\end{figure}

The second dependent variable aims to operationalize vaccination uptake through
vaccination rate information in the United States from January 2021 through
August 30 2021 \citep{cdcCOVID19Vaccination2021}. Data represents all vaccine partners including
jurisdictional partner clinics, retail pharmacies, long-term care facilities,
dialysis centers, Federal Emergency Management Agency and Health Resources and
Services Administration partner sites, and federal entity facilities.
Vaccination data is available for all US counties with the exception of parts of
California and Massachusetts, Hawaii, and Texas. In Texas and Hawaii, no county
level information is available, and California does not report the county of
residence for vaccinations when the county of residence has a population less
than 20,000 people. Finally, Massachusetts does not provide vaccination data for
Barnstable, Dukes, and Nantucket counties because of their small populations.
Therefore, my final county sample for vaccination rates is \emph{n} = 2819
(compared to the total population of 3,107 US counties). A list of included FIPS
are available upon request. Vaccination rates are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. I
sample every Monday between January 2021 through August 30, 2021 for 35 total
observations for each county.

@@

I utilize the likelihood of a friendship connection between counties
to create a county-level social network weighted by the probability of a tie.
Using this network and the two independent variables above, I examine how the
vaccination and stay-at-home rates of peer counties is contagious to the
ego-county. To do this, I first take the weighted average of each ego's network
signals using equation \eqref{eq:networksignal} where \texttt{x} denotes the
vaccination and stay-at-home rates of each alter county and \texttt{w} represents the
likelihood of a friendship connection between the ego county and each of their
alters, lagged by one week. This `signal' of norms gives us insight into the
coevolving contagion patterns of the new social norm being established. The
likelihood of friendship connection comes from the ``Social Connectedness Index''
\citep{Bailey2018, facebook20} which indexes the social links between geographies by
the likelihood of Facebook friends. It is an aggregated measure of Facebook
friendship connections between counties. It corresponds to ``the (relative)
probability that two arbitrary Facebook users across two geographies are friends
with each other.'' The data is available at various geographic aggregation
levels, such as U.S. zip codes or entire countries. However, data is only
available at one time point because Facebook has found that although there are
individual changes in friendship connections over time, the aggregate
statistical probabilities of friendship remain stable. This measure is based on
user-provided information and real-time location-service data gathered by
Facebook. Facebook data has been shown to be highly representative of the U.S.
population and Facebook friendship links largely represent real-world
friendships \citep{bailey_etal18, jones_etal13}. The data has been tested to show
how initial COVID-19 hotspots are related to subsequent virus spread to
non-hotspots, even after controlling for population density and geographic
distance \citep{Kuchler2020}.

@@

Figures \ref{fig:assortativity-vacc} and \ref{fig:discordancenetwork} provide
a visualization of both network signal and signal discordance for April 26, 2021
for 7 selected counties.\footnote{Figures in this paper were all created using ggplot2
\citep{wickham_etal, wickham11}, patchwork \citep{pedersen20}, and ggridges \citep{ggridges}}
While a county like Lake County, Ohio has low discordance meaning the signal of
vaccination rate is reinforced, Navajo County, Arizona receives very diverse
signals from their county-alters, negating any contagion effects. Figure
\ref{fig:discordancenetwork} explicitly outline how the weights and signals are
calculated for an ego with only 5 alters.

\begin{figure}
{\centering \includegraphics[width=0.7\linewidth]{figs/paper3/assortativity-vacc-1}}
\caption{How Signal Mean and Signal Assortativity are Measured}\label{fig:assortativity-vacc}
\end{figure}

\begin{figure}
\begin{center}\includegraphics[width=0.5\linewidth,]{figs/paper3/discordancenetwork-1} 
  \begin{equation}
    w = \begin{bmatrix}0.0000015\\0.0000023\\0.0000527\\0.0000966\\0.0001119\end{bmatrix},  
    x = \begin{bmatrix}29.9\\20.4\\27.7\\24.6\\24.3\end{bmatrix} \longrightarrow
    \begin{matrix} \bar{x}, signal = 25.083\\ s, discordance = 1.41 \end{matrix} 
  \end{equation}
  \caption{Visualization of how Network Signal and Discordance are Calculated}
  \label{fig:discordancenetwork}
  \end{center}
\end{figure}

@@

To estimate the concept of `real threat of infection' in figure
\ref{fig:norm-framework}, the models use county-level Covid-19 case rates with
data compiled by The New York Times \citeyearpar{covid_data}. It is widely acknowledged that there are biases in this data due to inconsistencies and availability in testing as well as
different community propensity to test \citep{gu22, cdc20a}. However, it is the best
measure we have of actual case rates. County data are scaled with a rolling mean
with a rolling window width of seven days to smooth out intra-week noise. Case
Rate is measured as number of cases per 100,000 population. Observations vary
from a minimum of 0 to a maximum of 1,565 across the two datasets.

@@

\input{tables/paper3/google-tab}
